---
title: "Assignment 1"
author: "Arnav Murulidhar"
date: "9/13/2020"
output: html_document
---

``` {r , message = FALSE}
library(ggplot2)
library(tidycensus)
library(tidyverse)
```

## Variables used: 
To create my data frame, I have filtered the database for residents of LA county who commute to work and live in a property of value.

1. Property value (continuous) : valp
2. Units in structure (categorical) : bld
3. Total person's income (continuous) : pincp
4. Means of transportation to work (categorical) : jwtr
5. Travel time to work (continuous) : jwmnp

```{r load person-level data, results = 'hide'}
person_data <- get_pums(variables = c("PUMA",
                                      "VALP", 
                                      "BLD", 
                                      "PINCP", 
                                      "JWTR", 
                                      "JWMNP"),
                        state = 06,
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(JWMNP > 0, VALP > 0, as.numeric(PUMA) == 03701) %>%
  select(VALP, BLD, PINCP, JWTR, JWMNP)
```
## Plot 1: Everything!
```{r everything scatter plot}
ggplot(person_data, 
       aes(x = VALP,
           y = PINCP,
           size = JWMNP,
           shape = JWTR,
           color = BLD)) +
  geom_point() +
  scale_size_continuous(name = "Tract population") +
  scale_x_continuous(name = "Property value",
                     breaks = seq(0, 10000000, by = 1000000)) +
  scale_y_continuous(name = "Income",
                     breaks = seq(0, 1000000, by = 100000)) +
  scale_color_discrete(name = "Housing type",
                       labels = c("One-family house detached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "One-family house attached",
                                  "5-9 Apartments",
                                  "3-4 Apartments",
                                  "50 or more Apartments")) +
  scale_shape(name = element_blank(), 
              labels = c("car, truck or van", 
                         "Bicyle",
                         "Bus or Trolley Bus",
                         "Railroad",
                         "Motorcycle",
                         "Walked",
                         "Other")) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Plot 2: Three variable scatter plot
```{r no-size scatter plot}
ggplot(person_data, 
       aes(x = VALP,
           y = PINCP,
           color = BLD,)) +
  geom_point() +
  scale_size_continuous(name = "Tract population") +
  scale_x_continuous(name = "Property value",
                     breaks = seq(0, 10000000, by = 1000000)) +
  scale_y_continuous(name = "Income",
                     breaks = seq(0, 1000000, by = 100000)) +
  scale_color_discrete(name = "Housing type",
                       labels = c("One-family house detached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "One-family house attached",
                                  "5-9 Apartments",
                                  "3-4 Apartments",
                                  "50 or more Apartments")) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Plot 3: Scatter with fill
```{r fill scatter plot}
ggplot(person_data, 
       aes(x = VALP,
           y = PINCP,
           color = BLD,
           fill = BLD)) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_size_continuous(name = "Tract population") +
  scale_x_continuous(name = "Property value",
                     breaks = seq(0, 10000000, by = 1000000),
                     labels = paste(seq(0, 10000000, by = 1000000), "k", sep = "")) +
  scale_y_continuous(name = "Income",
                     breaks = seq(0, 1000000, by = 100000),
                     labels = paste(seq(0, 1000000, by = 100000), "k", sep = "")) +
  scale_color_discrete(name = "Housing type",
                       labels = c("One-family house detached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "One-family house attached",
                                  "5-9 Apartments",
                                  "3-4 Apartments",
                                  "50 or more Apartments")) +
  scale_fill_discrete(name = "Housing type",
                       labels = c("One-family house detached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "One-family house attached",
                                  "5-9 Apartments",
                                  "3-4 Apartments",
                                  "50 or more Apartments")) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Plot 4: Two variable scatter
```{r two-var scatter plot}
ggplot(person_data, 
       aes(x = PINCP,
           y = JWMNP)) +
  geom_point() +
  stat_smooth(color = "black", linetype = 2, size = 0.5) +
  scale_x_continuous(name = "Income") +
  scale_y_continuous(name = "Commute time") +
  theme_bw()
```

## Plot 5: Two variable log scatter
```{r log scatter plot}
ggplot(person_data, 
       aes(x = PINCP,
           y = JWMNP)) +
  geom_point() +
  stat_smooth(color = "black", linetype = 2, size = 0.5) +
  scale_y_continuous(name = "commute time") +
  scale_x_continuous(name = "income",
                     trans = "log",
                     breaks = c(10, 100, 1000, 10000,100000,1000000),
                     labels = c("10", "100", "1,000", "10,000", "100,000", "1,000,000")) +
  theme_bw()
```

## Plot 6: Polar scatter
```{r polar scatter plot}
ggplot(person_data, 
       aes(x = PINCP,
           y = JWMNP)) +
  geom_point(alpha = 0.5, size = 0.5) +
  stat_smooth(color = "black", linetype = 2, size = 0.5) +
  scale_y_continuous(name = "commute time") +
  scale_x_continuous(name = "income",
                     trans = "log",
                     breaks = c(10, 100, 1000, 10000,100000,1000000),
                     labels = c("10", "100", "1,000", "10,000", "100,000", "1,000,000")) +
  coord_polar(theta = "x") +
  theme_bw()
```

## Plot 7: Categorical Variable scatter
```{r categ scatter plot}
ggplot(person_data, 
       aes(x = BLD,
           y = PINCP)) +
  geom_point() +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_y_continuous(name = "Income") +
  theme_bw() 
```

## Plot 8: Categorical jittered
```{r categ jitter scatter plot}
ggplot(person_data, 
       aes(x = BLD,
           y = PINCP,
           color = BLD)) +
  geom_point(position = "jitter",
             size = 0.75,
             alpha = 0.5) +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_y_continuous(name = "Income") +
  theme_bw() 
```

## Plot 9: Categorical box
```{r categ box plot}
ggplot(person_data, 
       aes(x = BLD,
           y = PINCP,
           color = BLD)) +
  geom_boxplot() +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_y_continuous(name = "Income") +
  theme_bw() 
```

## Plot 10: Categorical violin
```{r categ violin plot}
ggplot(person_data, 
       aes(x = BLD,
           y = PINCP,
           color = BLD)) +
  geom_violin() +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_y_continuous(name = "Income") +
  theme_bw() 
```

## Plot 11: Categorical scattered violin
```{r categ violin scatter plot}
ggplot(person_data, 
       aes(x = BLD,
           y = PINCP,
           color = BLD)) +
  geom_violin() +
  geom_point(position = "jitter",
             alpha = 0.5,
             size = 0.5) +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_y_continuous(name = "Income") +
  theme_bw() 
```

## Plot 12: Categorical bar
```{r categ bar chart}
ggplot(person_data, aes(x = BLD, fill = JWTR)) +
  geom_bar() +
  scale_y_continuous(name = "Number of people") +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_fill_discrete(name = element_blank(),
                        labels = c("car, truck or van", 
                         "Bicyle",
                         "Bus or Trolley Bus",
                         "Railroad",
                         "Motorcycle",
                         "Walked",
                         "Other")) +
  theme_bw() 
```

## Plot 13: Categorical dodged bar
```{r categ dodge bar chart}
ggplot(person_data, aes(x = BLD, fill = JWTR)) +
  geom_bar(position = "dodge") +
  scale_y_continuous(name = "Number of people") +
  scale_x_discrete(name = "Housing type",
                       labels = c("One-family\n house detached",
                                  "10-19\nApartments",
                                  "Mobile home\n or trailer",
                                  "One-family house\n attached",
                                  "5-9\n Apartments",
                                  "3-4\n Apartments",
                                  "50 \nor \nmore Apartments")) +
  scale_fill_discrete(name = element_blank(),
                        labels = c("car, truck or van", 
                         "Bicyle",
                         "Bus or Trolley Bus",
                         "Railroad",
                         "Motorcycle",
                         "Walked",
                         "Other")) +
  theme_bw() 
```

## Plot 14: Pie!
```{r pie chart}
pie_chart <- ggplot(person_data, aes(x = 1, fill = BLD)) +
  geom_bar(position = "fill") +
  coord_polar(theta = "y") +
  scale_y_continuous() +
  scale_fill_discrete(name = "Housing type",
                       labels = c("One-family house detached",
                                  "10-19 Apartments",
                                  "Mobile home or trailer",
                                  "One-family house attached",
                                  "5-9 Apartments",
                                  "3-4 Apartments",
                                  "50 or more Apartments")) +
  theme_void() 

pie_chart
```